They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-21 paper/paper.Rmd
re-reading, added all tables, by p-value adj, no p-values in tables

---
title: "They're Clutching up! Team Momentum in Round-Based Esports"
authors:
  - name: Tony ElHabr
    email: anthonyelhabr@gmail.com
    url: https://tonyelhabr.rbind.io
abstract: |
  My research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, I find evidence supprting the hypothesis that round win probability can be modeled as a constant across the series, although not at the naive 50%. Second, I examine post-streak round win probability, given the series outcome. I find no evidence that post-streak win rate is significantly different than expected for combinations of streak length, series length, and series winner, suggesting that the perception of momentum at the team level is deceiving.
bibliography: references.bib  
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

I adopt the terminology "series" to refer to what CoD SnD players typically call a "match", so as to emulate the terminology of playoff series in professional leagues like the National Basketball Association, National Hockey League, and Major League Baseball. A "game" or a "match" in such leagues is analogous to a "round" of CoD SnD.

@@

[^7]: Data: <https://linktr.ee/CDLArchive>. Author: <https://twitter.com/IOUTurtle>

@@

There have been a handful of studies of the distribution of games played in a series of a professional sport. Most assume a constant probability $\phi$ of a given team winning a game in the series, regardless of the series state. Mosteller [-@mosteller1952] observed that the American League had dominated the National League in Major League Baseball's (MLB) World Series matchups, implying that matchups should not modeled with $\phi = 0.5$. Mosteller proposed three approaches for identifying the optimal constant probability value of the stronger team in the World Series, finding $\phi \approx 0.65$. in each case: (1) solving for $p$ from the empirical average number of games won by the loser of the series, which he called the "method of moments" approach; (2) maximizing the likelihood that the sample would have been drawn from a population in which the probability of a team winning a game is constant across the series (i.e. maximum likelihood), and (3) minimizing the chi-square goodness of fit statistic for $\phi$.

Chance [-@chance2020] re-examines the constant probability notion in Major League Baseball's (MLB) World Series (1923--2018), the National Basketball Association's (NBA) Finals (1951--2018), and the National Hockey League's (NHL) Stanley Cup (1939--2018). Chance finds strong evidence against the null hypothesis of $\phi = 0.5$ in the MLB and NHL championship series when applying Mosteller's first and second methods. Chance goes on to outline a conditional probability framework (likelihood of winning a game given the series state) which can exactly explain the distribution of the number of games played.

Momentum, one of most discussed topics in sports analytics, goes hand-in-hand with a discussion of the nature of series outcomes.[^10] Two opposing fallacies are observed in the context of momentum: the "gambler's fallacy" (negative recency) and "hot hand fallacy" (positive recency). Per Ayton et al. [-@ayton2004], negative recency is "the belief that, for random events, runs of a particular outcome ... will be balanced by a tendency for the opposite outcome", while positive recency is the expectation of observing future results that match recent results.

@@

Despite the plethora of existing research on games played in a series and momentum in sports, these topics have yet to be investigated heavily in esports. Work has been done to examine in-round win probability in other FPS titles such as Counter-Strike [@xenopoulos2022] and Valorant [@derover2021], both of which are round-based like CoD SnD. However, research on round-level trends is sparse, perhaps because games like Counter-Strike and Valorant both have economic aspects that can create clear advantages on side in a given round, given how prior rounds played out.

Additionally, both Counter-Strike and Valorant have overtime rules and blocked offensive/defensive roles (i.e. playing either offense or defense for many consecutive rounds). On the other hand, teams in CoD SnD rotate sides every round, analogous to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series for professional sports like the MLB, NBA, and NHL.[^12] While theoretically one might be able to account for any kind of format, such as a 5-5-1, the rotation of team sides every round is convenient for convincing ourselves that rounds could reasonably be modeled as i.i.d. Bernoulli trials.

@@

@@

## Distribution of rounds played {#sec:method-rounds-played}

@@

## Momentum {#sec:method-momentum}

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq:pwr}{}{
p_0(\text{win} | r) = p^{w|r}_0 = \{
\begin{array}{rl}
\frac{m}{r}, & \text{team wins series} \\
\frac{r - m}{r}, & \text{team loses series}
\end{array}
}
\label{eq:pwr}
\end{equation}
```
using $m$ from Equation \ref{eq:m}.

As shown by MS with their Theorem 1, we should expect the proportion of rounds wins immediately following a streak of $k$ rounds wins for a series lasting $r$ rounds, $\hat{P}^{w|kr}_{MS}$, to be strictly less than $p^{w|r}$.[^15] Unfortunately, there does not exist a closed form representation for the expectation for $\hat{P}^{w|kr}_{MS}$ for $k > 1$. Nonetheless, one may recursively build a collection of dictionaries to compute the expectation analytically, or, more conveniently, one may run simulations to estimate the expected value.[^16]

[^15]: See Appendix E.1 of MS for details on the proof that follows.

[^16]: I've adapted the code from Vafa [-@vafa2017], which implements MS's framework [-@miller2018].

Although it would be convenient to emulate the hypothesis testing for difference of proportions performed by MS (and GVT), my context is fundamentally different from that of MS, who focus on longitudinal data in controlled settings.[^17] The number of trials is fixed in their experimental designs, but in CoD SnD, the number of rounds played is determined as a function of the max number of possible rounds ($s$) and whether or not the team wins the series, as shown in Equation \ref{eq:pwr}. The Bernoulli trial success probability, i.e. a round win in CoD SnD, is not independent of the opponent. Consequently, a statistical test of the difference in $\hat{P}^{w|kr}$ and $\hat{P}^{\ell|kr}$, as performed by MS to evaluate their hypothesis regarding post-streak success rates is not completely appropriate, although useful as a reference.

@@

One might consider another form of the expected proportion of rounds won immediately after a streak of $k$ round wins in a best-of-$s$ series given the length of the series ($r$ rounds), the "notional" proportion $\hat{P}^{w|kr}_0$. The Bernoulli round win probability underlying $\hat{P}^{w|kr}_0$ is

```{=tex}
\begin{equation}\protect\hypertarget{eq:pwkr}{}{
p_0(\text{win} | k, r) = p^{w|kr}_0 =\{
\begin{array}{rl}
\frac{m - k}{i - k}, & \text{team wins series} \\
\frac{s - m - k}{i - k}, & \text{team loses series}
\end{array}
}
\label{eq:pwkr}
\end{equation}
```
using $m$ from Equation \ref{eq:m}. We can perform a binomial test to test the null hypothesis $H_0 : \omega = \omega_0$ for the observed probability of success $\omega$ and a user-specified $\omega_0$, where $0 \leq \omega_0 \leq 1$.[^18]

@@

If there are $r_+$ observed successes in a sample of $r$ trials and we expect that there should be $r * \omega_0$, the probability of arriving at this expected number of successes is

```{=tex}
\begin{equation}\protect\hypertarget{eq:binom}{}{
\Pr(r_+) = {\binom {r}{r_+}} \omega^{r_+}(1-\omega)^{r-r+}.
}\label{eq:binom}\end{equation}
```
Plugging in the notional probability $p^{w|kr}_0$ for the null $\omega_0$ and the observed probability $p^{w|kr} = N^{w|kr} \times P^{w|kr}$, for the observed $\omega$, where $N^{w|kr}$ is the observed number of instances where a teams wins the round immediately following a $k$-round winning streak in a series lasting $r$ rounds, we can evaluate the null hypothesis that the observed probability is equal to the notional probability. If we reject this null hypothesis, then we might consider team momentum, represented by post-streak success, plausible.

One can perform the same binomial test for the MS probability adjusting for streak-selection bias, $p^{w|kr}_{MS}$. Given the caveats mentioned before, the results should be heeded with caution.

Stepping back from the one-sided perspective of a single team's probability of winning and their streakiness, one can attempt to detect the "hot hand" more broadly with the Wald-Wolfowitz runs test, a common tool used in the field of quality control. Under the null hypothesis, the number of runs in a sequence of $r$ trials is a random variable with the following mean $\mu$ and variance $\sigma^2$.

```{=tex}
\begin{equation}\protect\hypertarget{eq:ww}{}{
\mu = \frac{2r_+r_-}{r} + 1, \sigma^2 = \frac{(\mu-1)(\mu-2)}{r-1}
}\label{eq:ww}\end{equation}
```
One might subset observed series sequences to those that violate the run test null hypothesis and perform a proportions test where the null is that the observe rate of such series is equal to the frequency that we should expect, given a choice for the constant round win probability $\phi$. If the

@@

## Distribution of rounds played {#sec:results-rounds-played}

@@

## Momentum {#sec:results-momentum}

@@

cod_round_win_prop_after_k_wins <- tibble(win_streak = 2L:5L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins)
  ) |>
  unnest(data)
cod_round_win_prop_after_k_wins |> filter(win_round, notional_p_value < 0.05)
cod_round_win_prop_after_k_wins |> filter(win_round, ms_p_value < 0.05)
# bonferroni correction?
cod_round_win_prop_after_k_wins |> 
  mutate(
    n = n(),
    p2 = notional_p_value |> p.adjust(),
    p3 = ms_p_value |> p.adjust()
  ) |> 
  filter(win_streak == 3, win_round)
```

```{r}
#| label: cod_round_win_prop_after_k_wins-table
cod_round_win_prop_after_k_wins |> 
  filter(win_streak == 2L) |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        TRUE ~ sprintf('%.2f', .x)
      )
    ),
    prop_lab = scales::percent(prop, accuracy = 0.1),
    notional_lab = sprintf('%.1f%% (%s)', round(100 * notional_prop, 1), notional_p_value),
    ms_lab = sprintf('%.1f%% (%s)', round(100 * ms_prop, 1), ms_p_value)
  ) |>
  arrange(n_rounds, win_series, win_round) |> 
  transmute(
    n_rounds,
    across(win_series, ~ifelse(.x, 'yes', 'no')),
    win_round,
    total,
    prop_lab,
    notional_lab,
    ms_lab
  ) |>
  group_split(win_round) |> 
  walk(
    function(df) {
      message(sprintf('win_round=%s, win_series=%s', df$win_round[[1]], df$win_series[[1]]))
      df <- df |> 
        select(
          -c(win_round)
        ) |> 
        gt::gt() |> 
        gt::cols_label(
          n_rounds = 'r',
          win_series = 'Win series?',
          total = 'N',
          prop_lab = 'O',
          notional_lab = 'a',
          ms_lab = 'b'
        ) |> 
        gt::as_latex() |> 
        cat()
    }
  )

@@

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on streaks of three round wins.[^20] Table \ref{tbl:cod-pw3r-pl3r} compares the observed round win rate $P^{w,kr}$ following streaks of $k=3$ round wins given the series outcome ($r$ rounds and winner) with the notional and MS expected proportions, $\hat{P}^{w|kr}_0$ and $\hat{P}^{w|kr}_{MS}$ respectively.

```{=tex}
\begin{longtable}{rcrrrr}
\caption{Observed count and proportion of round wins immediately following round win streak of $k=3$, $N^{w|kr}$ and $P^{w|kr}$ respectively, given the length of the series ($r$ rounds) and the series winner. Additionally, the notional and MS expected proportions, $\hat{P}^{w|kr}_0$ and $\hat{P}^{w|kr}_{MS}$ respectively, and the $p$-values of binomial tests comparing the underlying notional and MS round win probabilities, $p^{w|kr}_0$ and $p^{w|kr}_{MS}$, to the probability implied by the observed proportion $p^{w|kr} = N^{w|kr} \times P^{w|kr}$.}
\label{tbl:cod-pw3r-pl3r} \\
\toprule
$r$ & \text{Win series?} & $N^{w|kr}$ & $P^{w|kr}$ & $\hat{P}^{w|kr}_0$ ($p$-value) & $\hat{P}^{w|kr}_{MS}$ ($p$-value) \\ 
\midrule
7 & yes & 209 & 74.6\% & 75.0\% (0.94) & 75.7\% (0.69) \\ 
8 & yes & 209 & 62.2\% & 60.0\% (0.53) & 61.9\% (0.94) \\ 
9 & yes & 193 & 51.8\% & 50.0\% (0.67) & 52.7\% (0.83) \\ 
10 & no & 60 & 13.3\% & 14.3\% (1.00) & 26.7\% (0.02) \\ 
10 & yes & 151 & 43.7\% & 42.9\% (0.87) & 44.7\% (0.87) \\ 
11 & no & 129 & 24.0\% & 25.0\% (0.84) & 30.8\% (0.11) \\ 
11 & yes & 150 & 40.0\% & 37.5\% (0.56) & 38.8\% (0.80) \\

\bottomrule
\end{longtable}
```
We observe that, with the exception of $\hat{P}^{w|kr}_{MS}$ when $r = 10$ and the eventual series loser is the team that wins after a streak of three round wins[^21], all binomial test $p$-values are greater than the typical $\alpha = 0.05$ threshold, implying that we cannot reject the null hypothesis that the expected notional and MS post-streak round win rates are different than the observed proportion. As noted before, the MS expectations may be unreliable for CoD SnD, so one is inclined to prefer the results of the notional binomial tests.

[^21]: This implies that the opponent, th eventual series winner, goes on to win the series 6-4 after going down 0-4. This is the only scenario in which one team wins four rounds consecutively but loses a best-of-11 series in 10 rounds.

When performing the same tests for streaks of two, four, and five, there is no case in which I can reject the binomial null hypothesis for the expected notional probability $p^{w|kr}_0$. The null hypothesis can be rejected for expected MS probability $p^{w|kr}_{MS}$ when $k = 4, r = [7, 8, 10]$ and $k = 5, r = [7, 8]$. Nonetheless, the usefulness of $p^{w|kr}_{MS}$ is questionable, as described before, and the sample sizes are smaller for these conditions (all under 20 instances, with the exception of $k = 4, r = [7, 8]$ with 70 and 39 observations respectively).

>>>>

---
title: "They're Clutching up! Team Momentum in Round-Based Esports"
authors:
  - name: Tony ElHabr
    email: anthonyelhabr@gmail.com
    url: https://tonyelhabr.rbind.io
abstract: |
  My research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, I find evidence supprting the hypothesis that round win probability can be modeled as a constant across the series, although not at the naive 50%. Second, I examine post-streak round win probability, given the series outcome. I find no evidence that post-streak win rate is significantly different than expected for combinations of streak length, series length, and series winner, suggesting that the perception of momentum at the team level is just that, perception.
bibliography: references.bib  
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

I adopt the terminology "series" to refer to what CoD SnD players typically call a "match", so as to emulate the terminology of playoff series in professional leagues like the National Basketball Association, National Hockey League, and Major League Baseball. A "game" or a "match" in such leagues is analogous to a "round" of CoD SnD. Further, it is analogous to a trial in statistical analysis of Bernoulli events.

@@

[^7]: Raw data: <https://linktr.ee/CDLArchive>. Author: <https://twitter.com/IOUTurtle>. Processed data can be found at <https://github.com/tonyelhabr/fps_round_games/blob/master/data/cod_rounds.csv>.

@@

There have been a handful of studies of the distribution of games played in a series of a professional sport. Most assume a constant probability $\phi$ of a given team winning a game in the series, regardless of the series state. Mosteller [-@mosteller1952] observed that the American League had dominated the National League in Major League Baseball's (MLB) World Series matchups, implying that games should not be modeled with a constant (null) $\phi_0 = 0.5$. Mosteller proposed three approaches for identifying the optimal constant probability value of the stronger team in the World Series, finding $\phi \approx 0.65$ in each case: (1) solving for $p$ from the empirical average number of games won by the loser of the series, which he called the "method of moments" approach; (2) maximizing the likelihood that the sample would have been drawn from a population in which the probability of a team winning a game is constant across the series (i.e. maximum likelihood), and (3) minimizing the chi-square goodness of fit statistic $\chi^2$ for $\phi$.

Chance [-@chance2020] re-examines the constant probability notion in Major League Baseball's (MLB) World Series (1923--2018), the National Basketball Association's (NBA) Finals (1951--2018), and the National Hockey League's (NHL) Stanley Cup (1939--2018). Chance finds strong evidence against the null hypothesis of $\phi_0 = 0.5$ in the MLB and NHL championship series when applying Mosteller's first and second methods. Chance goes on to outline a conditional probability framework (likelihood of winning a game given the series state) which can exactly explain the distribution of the number of games played.

Momentum, one of most discussed topics in sports analytics, goes hand-in-hand with a discussion of the nature of series outcomes.[^10] Two opposing fallacies are observed in the context of momentum: the "gambler's fallacy" (negative recency) and the "hot hand fallacy" (positive recency). Per Ayton et al. [-@ayton2004], negative recency is "the belief that, for random events, runs of a particular outcome ... will be balanced by a tendency for the opposite outcome", while positive recency is the expectation of observing future results that match recent results.

@@

Other research has borrowed a common approach from the field of quality control, looking at unlikely sequences of events with the Wald-Wolfowitz runs test [@peel2015], [@steeger2021]. Peel and Clauset find no evidence for unlikely sequences of scoring events in the NHL, College Football (CFB), and National Football League (NFL), but do in the NBA. As a check on their entropy approach to momentum identification, Steeger et al. found three NHL teams who sequence of wins in the 2018-2019 regular season violated the Wald-Wolfowitz null hypothesis, although these teams did not match the teams identified with their primary approach.

## My contribution

Despite the plethora of existing research on games played in a series and momentum in sports, these topics have yet to be investigated heavily, if at all, in esports. Work has been done to examine intra-round win probability in other FPS titles such as Counter-Strike [@xenopoulos2022] and Valorant [@derover2021], both of which are round-based like CoD SnD. However, considering Counter-Strike and Valorant specifically, research on round-level trends seems non-existant, perhaps for one of many reasons:

1.  Both have economic aspects that can create clear advantages for one side in a given round, given how prior rounds played out. CoD has no such equivalent, except for perhaps "kill streaks", which infrequently play a role.
2.  Teams play either offense or defense for many consecutive rounds. (The former has a 14-14-1 format and the latter uses a 12-12-1 format.) On the other hand, teams in CoD SnD rotate roles every round, analogous to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series for professional sports like the MLB, NBA, and NHL.[^12] While theoretically one might be able to account for any kind of format, such as a 5-5-1, the rotation of team sides every round is convenient for convincing ourselves that rounds could reasonably be modeled as i.i.d. Bernoulli trials.
3.  Both Counter-Strike and Valorant have overtime rules---team must win by two rounds---which can make end-of-series sequences difficult to model. CoD SnD does not have overtime rules.

@@

To my knowledge, there is no existing public statistical research on CoD SnD format, beyond descriptive analysis on social media. While intra-round trends may be more directly applicable to teams looking for an advantage on their competition, broader investigation on a common concept like the hot-hand phenomenon in a new field---particularly one that is less subjected to factors that may be difficult to control for, e.g. weather---should hopefully be useful as a precedent or supplement for future researchers.

@@

## Distribution of rounds played

@@

## Momentum

### MS post-streak probability

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq:pwr}{}{
p_0(\text{win} | r) = p^{+|r}_0 = \{
\begin{array}{cl}
\frac{m}{r}, & \text{team wins series} \\
\frac{r - m}{r}, & \text{team loses series}
\end{array}
}
\label{eq:pwr}
\end{equation}
```
using $m$ from Equation \ref{eq:m}.

As shown by MS with their Theorem 1, we should expect the proportion of rounds wins immediately following a streak of $k$ rounds wins for a series lasting $r$ rounds, $\hat{P}^{+|kr}_{MS}$, to be strictly less than $p^{+|r}$.[^15] Unfortunately, there does not exist a closed form representation for the expected value of $\hat{P}^{+|kr}_{MS}$ for $k > 1$. Nonetheless, one may recursively build a collection of dictionaries to compute the expectation analytically, or, more conveniently, one may run simulations to estimate the expected value, as I choose to do.[^16]

[^15]: See Appendix E.1 of MS for the proof.

[^16]: I have adapted the code from Vafa [-@vafa2017], which implements MS's framework [-@miller2018].

Although it would be convenient to emulate the hypothesis testing for difference of proportions performed by MS (and GVT), my context is fundamentally different from that of MS, who focus on longitudinal data in controlled settings.[^17] The number of trials is fixed in their experimental designs, but in CoD SnD, the number of rounds played is determined as a function of the max number of possible rounds ($s$) and whether or not the team wins the series. The Bernoulli trial success probability, i.e. a round win in CoD SnD, is not independent of the opponent. Consequently, a statistical test of the difference in $\hat{P}^{+|kr}$ and $\hat{P}^{-|kr}$, as performed by MS to evaluate their hypothesis regarding post-streak success rates, is not completely appropriate, although useful as a reference.

@@

### "Notional" post-streak probability

One might consider another form of the expected proportion of rounds won immediately after a streak of $k$ round wins in a best-of-$s$ series given the length of the series ($r$ rounds), the "notional" proportion $\hat{P}^{+|kr}_0$. The Bernoulli round win probability underlying $\hat{P}^{+|kr}_0$ is

```{=tex}
\begin{equation}\protect\hypertarget{eq:pwkr}{}{
p_0(\text{win} | k, r) = p^{+|kr}_0 =\{
\begin{array}{cl}
\frac{m - k}{r - k}, & \text{team wins series} \\
\frac{s - m - k}{r - k}, & \text{team loses series}
\end{array}
}
\label{eq:pwkr}
\end{equation}
```
using $m$ from Equation \ref{eq:m}.

We can perform a binomial test to evaluate the null hypothesis $\omega = \omega_0$ for the observed probability of success $\omega$ and a user-specified $\omega_0$, where $0 \leq \omega_0 \leq 1$.[^18] If there are $r^+$ observed successes in a sample of $r$ trials and we expect that there should be $r * \omega_0$, the probability of arriving at this expected number of successes is

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq:binom}{}{
\Pr(r^w) = {\binom {r}{r^+}} \omega^{r^+}(1-\omega)^{r-r^+}.
}\label{eq:binom}\end{equation}
```
Plugging in the notional proportion $\hat{p}^{+|kr}_0$ for the null $\omega_0$ and the observed proportion $P^{+|kr}$ for $\omega$ (treating proportions as probabilities), we can evaluate the null hypothesis that the observed probability is equal to the notional probability. If we can reject this null hypothesis, then we might consider team momentum, represented by post-streak success, plausible.

One can perform the same binomial test for the MS's streak-selection-adjusted proportion, $\hat{P}^{+|kr}_{MS}$. FOOTY: Given the caveats mentioned before, the results should be heeded with caution.

We can further decompose Equation \ref{eq:pwkr} by the round $i$ ($i \leq r$) in which the streak of length $k$ carries into.

```{=tex}
\begin{equation}\protect\hypertarget{eq:pwkri}{}{
p_0(\text{win} | k, r, i) = p^{+|kri}_0 =\{
\begin{array}{cl}
\frac{m - k}{r - k}, & \text{team wins series}, i \neq r \\
1, & \text{team wins series}, i = r \\
\frac{s - m - k}{r - k}, & \text{team loses series}, i \neq r \\
0, & \text{team loses series}, i = r
\end{array}
}
\label{eq:pwkri}
\end{equation}
```
Again, one might apply a binomial test to evaluate the hypothesis that the expected proportions, both $\hat{P}^{+|kri}_0$ and $\hat{P}^{+|kri}_{MS}$, are equal to the observed proportion $P^{+|kr}$.

### Wald-Wolfowitz runs test

Stepping back from the one-sided perspective of a single team's round win probability when streaking, one can attempt to detect the hot hand phenomenon with the Wald-Wolfowitz runs test. Under the null hypothesis, the number of runs in a sequence of $r$ trials is a random variable, which can take on values $r^+$ and $r^-$ for success and failure respectively, with the following mean $\mu$ and variance $\sigma^2$.

```{=tex}
\begin{equation}\protect\hypertarget{eq:ww}{}{
\mu = \frac{2r^{+}r^{-}}{r} + 1, \sigma^2 = \frac{(\mu-1)(\mu-2)}{r-1}
}\label{eq:ww}\end{equation}
```
One might subset observed series sequences to those that violate the run test null hypothesis and perform a test of proportions, where the null is that the observed rate of such series is equal to the frequency that we should expect, given a choice for the constant round win probability $\phi$. If we can reject the null, then we can argue in favor of the hot hand.

@@

## Distribution of rounds played

@@

## Momentum

@@

cod_round_win_prop_after_k_wins <- tibble(win_streak = 2L:5L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins)
  ) |>
  unnest(data)
cod_round_win_prop_after_k_wins |> filter(win_round, notional_p_value < 0.05)
cod_round_win_prop_after_k_wins |> filter(win_round, ms_p_value < 0.05)
```

```{r}
#| label: cod_round_win_prop_after_k_wins-table
cod_round_win_prop_after_k_wins |> 
  mutate(
    orig_notional_p_value = notional_p_value,
    orig_ms_p_value = ms_p_value
  ) |> 
  # group_by(win_streak) |> 
  mutate(
    across(c(notional_p_value, ms_p_value), ~p.adjust(.x, method = 'BY'))
  ) |> 
  # ungroup() |>
  filter(win_streak == 3L) |> 
  # filter(ms_p_value < 0.05)
  mutate(
    prop_lab = scales::percent(prop, accuracy = 0.1),
    notional_lab = sprintf('%.1f%%', 100 * notional_prop),
    ms_lab = sprintf('%.1f%%', 100 * ms_prop)
  ) |>
  arrange(n_rounds, win_series, win_round) |> 
  filter(win_round) |> 
  transmute(
    n_rounds,
    across(win_series, ~ifelse(.x, 'yes', 'no')),
    n,
    total,
    prop_lab,
    notional_lab,
    ms_lab
  ) |> 
  gt::gt() |> 
  gt::cols_label(
    n_rounds = 'r',
    win_series = 'Win series?',
    n = 'r+',
    total = 'N',
    prop_lab = 'O',
    notional_lab = 'a',
    ms_lab = 'b'
  ) |> 
  gt::as_latex() |> 
  cat()
```

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on streaks of three round wins.[^20] Table \ref{tbl:cod-pw3r-pl3r} compares the observed round win rate $P^{w,kr}$ following streaks of $k=3$ round wins given the series outcome ($r$ rounds and winner) with the notional and MS expected proportions, $\hat{P}^{+|kr}_0$ and $\hat{P}^{+|kr}_{MS}$ respectively.

```{=tex}
\begin{longtable}{rcrrrrr}
\caption{Given the round win streak $k$, the length of the series ($r$ rounds), and the series winner, the observed count of rounds wins, $r^{+|kr}$, and proportion of round wins, $P^{+|kr}$, among $N^{kr}$ instances where a team could win after the streak $k=3$ ($P^{+|kr} = \frac{r^{+|kr}}{N^{kr}}$). Additionally, the notional and MS expected proportions, $\hat{P}^{+|kr}_0$ and $\hat{P}^{+|kr}_{MS}$ respectively.}
\label{tbl:cod-pw3r-pl3r} \\

\toprule
$r$ & \text{Win series?} & $r^{+|kr}$ & $N^{kr}$ & $P^{+|kr}$ & $\hat{P}^{+|kr}_0$ & $\hat{P}^{+|kr}_{MS}$ \\ 
\midrule

7 & yes & 156 & 209 & 74.6\% & 75.0\% & 75.7\% \\ 
8 & yes & 130 & 209 & 62.2\% & 60.0\% & 61.9\% \\ 
9 & yes & 100 & 193 & 51.8\% & 50.0\% & 52.7\% \\ 
10 & no & 8 & 60 & 13.3\% & 14.3\% & 26.7\% \\ 
10 & yes & 66 & 151 & 43.7\% & 42.9\% & 44.7\% \\ 
11 & no & 31 & 129 & 24.0\% & 25.0\% & 30.8\% \\ 
11 & yes & 60 & 150 & 40.0\% & 37.5\% & 38.8\% \\

\bottomrule
\end{longtable}
```
With the exception of $\hat{P}^{+|kr}_{MS}$ when $r = 10$ and the eventual series loser is the team that wins after a streak of three round wins[^21], all un-adjusted binomial test $p$-values are greater than the typical $\alpha = 0.05$ threshold, implying that we cannot reject the null hypothesis that the expected notional and MS post-streak round win rates are different than the observed proportion. Adjusting the $p$-values with the Benjamini & Yekutieli (BY) correction [-@benjamini2001], the null hypothesis cannot be rejected for any case.[^22]

[^21]: This implies that the opponent, the eventual series winner, goes on to win the series 6-4 after going down 0-4. This is the only scenario in which one team wins four rounds consecutively but loses a best-of-11 series in 10 rounds.

[^22]: As noted before, the MS expectations may be unreliable for CoD SnD, so one is inclined to prefer the results of the notional binomial tests.

When performing the same tests for streaks of two, four, and five, there is no case in which I can reject the binomial null hypothesis for the expected notional probability $p^{+|kr}_0$. The null hypothesis can only be rejected for the expected MS probability $p^{+|kr}_{MS}$ when $k = 5, r = 7$ after applying the BY p-value correction.[^23]

[^23]: At 36, the sample size for \$k = 5, r = 7\$ is not as large as most other combinations of \$k\$ and \$r\$.

```{r}
#| label: cod_round_win_prop_after_k_wins_i_rounds
summarize_cod_win_prop_after_k_wins_i_rounds <- function(k) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    round = 3L:11L,
    win_series = c(TRUE, FALSE),
    win_round = c(TRUE, FALSE)
  ) |> 
    filter(round >= (!!k + 1L), n_rounds > round) |> 
    compute_probs_nx_k(k)

init_cod_round_win_prop_after_k_wins_i_rounds <- tibble(win_streak = 3L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins_i_rounds)
  ) |>
  unnest(data) |> 
  select(-starts_with('ms'))

cod_round_win_prop_after_k_wins_i_rounds <- init_cod_round_win_prop_after_k_wins_i_rounds |>
  group_by(win_streak, win_round, n_rounds, round) |> 
  mutate(
    total = sum(n),
    prop = n / total
  ) |> 
  ungroup() |> 
  mutate(
    notional_p_value = vectorized_binom_test(n, total, p = notional_prop)$p.value |> round(2)
  )

# ## example of why we needed to recalculate prop... prop and notional_prop should sum to 1, grouped by win_streak and win_round
# cod_round_win_prop_after_k_wins_i_rounds |> 
#   filter(win_streak == 3, n_rounds == 11, round == 10)
# 
# ## 5-3?
# cod_round_win_prop_after_k_wins_i_rounds |> 
#   filter(win_streak == 2) |> 
#   filter(n_rounds == 11, round == 11)
```

```{r}
#| label: cod_round_win_prop_after_k_wins_i_rounds-table

cod_round_win_prop_after_k_wins_i_rounds |> 
  mutate(
    across(notional_p_value, p.adjust)
  ) |>
  filter(
    win_round,
    # round >= (n_rounds - 3)
    n_rounds >= 10,
    win_streak == 3
  ) |>  
  filter(total >= 10) |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        TRUE ~ sprintf('%.2f', .x)
      )
    ),
    prop_lab = scales::percent(prop, accuracy = 0.1),
    notional_lab = scales::percent(notional_prop, accuracy = 0.1)
  ) |>
  arrange(n_rounds, round, win_series, win_round) |> 
  transmute(
    n_rounds,
    round,
    across(win_series, ~ifelse(.x, 'yes', 'no')),
    n,
    total,
    prop_lab,
    notional_lab
  ) |> 
  gt::gt() |> 
  gt::cols_label(
    n_rounds = 'r',
    round = 'i',
    win_series = 'Win series?',
    n = 'r+',
    total = 'N',
    prop_lab = 'O',
    notional_lab = 'a'
  ) |> 
  gt::as_latex() |> 
  cat()

@@

Now let us explicitly consider the round, $i$. As shown in Table \ref{tbl:cod-pw3ri-pl3ri}, the notional win rate in the round $i$ immediately following a streak of $k$ round wins in a series lasting $r$ rounds, $\hat{P}{+|kri}_0$, is statistically different than the observed $P^{+|kri}$ in several cases, notably when $r = i$ for each of $r \in [9, 10, 11]$.[^24] Interestingly, the observed rate is greater than the expected rate in each case where we can reject the null hypothesis, whereas the observed rate is less than the expected rate in all other cases, as well in a strong majority of cases not shown.

[^24]: The null can also be rejected for $r \in [7, 8]$, although these are not shown, and the sample sizes are smaller.

```{=tex}
\begin{longtable}{rrcrrrr}
\caption{Given the round $i$, the round win streak $k$, the length of the series ($r$ rounds), and the series winner, the observed count of rounds wins, $r^{+|kri}$, and proportion of round wins, $P^{+|kri}$, among $N^{kri}$ instances where a team could win after the streak $k=3$ ($P^{+|kri} = \frac{r^{+|kri}}{N^{kri}}$). Additionally, the notional proportion, $\hat{P}^{+|kri}_0$, and the $p$-value of a binomial test comparing $\hat{P}^{+|kri}_0$ to the observed proportion $P^{+|kri}$. Table restricted to $r \in [10, 11]$ and $i < r$ and $N^{kri} > 10$ for brevity.}
\label{tbl:cod-pw3ri-pl3ri} \\
\toprule
$r$ & $i$ & \text{Win series?} & $r^{+|kri}$ & $N^{kri}$ & $P^{+|kri}$ & $\hat{P}^{+|kri}_0$\\ 
\midrule

10 & 7 & no & 2 & 13 & 15.4\% & 14.3\% \\ 
10 & 7 & yes & 11 & 13 & 84.6\% & 42.9\% \\ 
10 & 8 & no & 1 & 10 & 10.0\% & 14.3\% \\ 
10 & 8 & yes & 9 & 10 & 90.0\% & 42.9\% \\ 
11 & 5 & no & 9 & 13 & 69.2\% & 25.0\% \\ 
11 & 5 & yes & 4 & 13 & 30.8\% & 37.5\% \\ 
11 & 6 & no & 4 & 13 & 30.8\% & 25.0\% \\ 
11 & 6 & yes & 9 & 13 & 69.2\% & 37.5\% \\ 
11 & 7 & no & 3 & 10 & 30.0\% & 25.0\% \\ 
11 & 7 & yes & 7 & 10 & 70.0\% & 37.5\% \\ 
11 & 9 & no & 5 & 15 & 33.3\% & 25.0\% \\ 
11 & 9 & yes & 10 & 15 & 66.7\% & 37.5\% \\ 
11 & 10 & no & 5 & 10 & 50.0\% & 25.0\% \\ 
11 & 10 & yes & 5 & 10 & 50.0\% & 37.5\% \\

\bottomrule
\end{longtable}
```
```{r}
#| label: cod_round_arrangements
library(randtests)
simulate_cod_round_arrangements <- function(p = 0.5, n_sims = 10000, seed = 42) {
  n_rounds_max <- 11
  set.seed(seed)
  w <- sample(c(0, 1), size = n_rounds_max * n_sims, replace = TRUE, prob = c(1 - p, p))
  m <- matrix(w, nrow = n_sims, ncol = n_rounds_max)
  # data.frame(x)
  df <- as_tibble(m)
  names(df) <- sprintf('%d', 1:n_rounds_max)
  df$i <- 1:nrow(df)
